---
title: "Rare variants analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library(knitr))
suppressMessages(library(data.table))
suppressMessages(library(tidyverse))
suppressMessages(library(qvalue))

outlier <- function (x) {
  IQR <- quantile(x)[[4]] - quantile(x)[[2]]
  low <- quantile(x)[[2]] - 1.5*IQR
  high <- quantile(x)[[4]] + 1.5*IQR
  out_val <- x[x < low | x > high]
  return(out_val)
}
```


```{r loading_data, echo=F}
res1 <- fread('../../../output_data/OUTPUT/210719/regenieRes/MNM/burden.res.common.1.perc_Covid_resistance.regenie') %>% 
  na.omit() %>%
  arrange(Pval,Alt) %>%
  mutate(Name = gsub("\\..*","",Name))

res1 %>% filter(Pval < 0.01) %>% 
  write.table('res1_sig.txt',
              col.names = T,
              row.names = F,
              sep='\t',
              quote=F)

res01 <- fread('../../../output_data/OUTPUT/210719/regenieRes/MNM/burden.res.common.0.1.perc_Covid_resistance.regenie') %>% 
  na.omit() %>%
  arrange(Pval) %>%
  mutate(Name = gsub("\\..*","",Name))

res01 %>% filter(Pval < 0.01) %>% 
  write.table('res01_sig.txt',
              col.names = T,
              row.names = F,
              sep='\t',
              quote=F)

```



### 1. Rare variants with MAF < 1%

```{r res_1_q_manhattan_plot, echo=F}
res1_man <- res1 %>% 
  mutate(Padj = 0.01/length(unique(res1$Name))) %>%
  relocate(Padj, .after=Pval) %>%
  arrange(Padj)

res1_man$chr_col <- NA
res1_man$chr_col <- ifelse(res1_man$Chr %% 2 == 0, 'even','odd')
res1_man$chr_col <- ifelse(res1_man$Pval <= 0.01 & grepl(res1_man$Alt, pattern = 'M3') == T, 's',res1_man$chr_col)

res1_man$chr_col <- ifelse(res1_man$Pval <= 0.01 & grepl(res1_man$Alt, pattern = 'M1') == T, 's',res1_man$chr_col)


chr.group <- c(even = '#27384A',odd = '#48C095',s='#BC0020')


res1_man %>%
  filter(is.na(Pval) == F) %>%
  ggplot(aes(x=Chr,y=-log10(Pval))) +
  geom_jitter(aes(col=chr_col)) +
  scale_color_manual(values = chr.group) +
  theme_classic() +
  theme(legend.position = 'none') +
  geom_hline(yintercept = c(-log10(0.01)),
              col='#BC0020',
              linetype='dashed') +
  scale_x_continuous(labels = c(1:22,'X'),breaks = 1:23) +
  xlab('Chromosomes')
```

### 2. Rare variants with MAF < 0.1%

```{r res_01_q_manhattan_plot, echo=F}
res01_man <- res01 %>% mutate(Qval = qvalue(Pval)$qvalue) %>%
  relocate(Qval, .after=Pval)

res01_man$chr_col <- NA
res01_man$chr_col <- ifelse(res01_man$Chr %% 2 == 0, 'even','odd')
res01_man$chr_col <- ifelse(res01_man$Pval <= 0.01 & grepl(res01_man$Alt, pattern = 'M3') == T, 's',res01_man$chr_col)

res01_man$chr_col <- ifelse(res01_man$Pval <= 0.01 & grepl(res01_man$Alt, pattern = 'M1') == T, 's',res01_man$chr_col)


chr.group <- c(even = '#27384A',odd = '#48C095',s='#BC0020')

res01_man %>% 
    filter(is.na(Pval) == F) %>%
  ggplot(aes(x=Chr,y=-log10(Pval))) +
    geom_jitter(aes(col=chr_col)) +
  scale_color_manual(values = chr.group) +
  theme_classic() +
  theme(legend.position = 'none') +
  geom_hline(yintercept = c(-log10(0.01)),
              col='#BC0020',
              linetype='dashed') +
  scale_x_continuous(labels = c(1:22,'X'),breaks = 1:23) +
  xlab('Chromosomes')
```

